Rate Transfers with Polyhedral Surrogate Losses /\\ Surrogate Regret Bounds for Polyhedral Losses

### 2021-05-20 rates-neurips-21/rates.tex
tweaks and edits in upper bound of rates

\section{Upper bound: polyhedral surrogates}
In this section, we develop the proof of Theorem \ref{thm:main-upper}.
At a high level, the proof has two parts:
\begin{enumerate}
  \item Fix $p \in \simplex$.
        Prove that $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$ for some constant $\alpha_p$.
  \item Apply this result to a carefully chosen finite subset of $\simplex$. Argue that the maximum over the $\alpha_p$'s in this finite set gives the constant factor for the transfer rate.
\end{enumerate}
We view both parts as relatively surprising.
The first part will actually hold for any calibrated surrogate.
The second part relies on a crucial fact about polyhedral losses $L$: their properties $\Gamma$ cover the whole simplex with a finite number of polyhedral level sets.
In other words, although the prediction space $\reals^d$ is infinite, one can pick out just a finite set of prediction points that can act as minimizers of expected loss.

\paragraph{Part 1.}
\begin{lemma} \label{lemma:fixed-p}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss and suppose the surrogate $L: \reals^d \to \reals_+^{\Y}$ and link $\psi: \reals^d \to \R$ are calibrated for $\ell$.
  Then for any $p \in \simplex$, there exists $\alpha_p \geq 0$ such that, for all $u \in \reals^d$,
    \[ R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p) . \]
\end{lemma}
\begin{proof}
  Fix $p \in \simplex$.
  Let $C_p = \max_{r \in \R} R_{\ell}(r,p)$.
  The maximum exists because $\ell$ is discrete, i.e. $\R$ is finite.
  Meanwhile, recall that, when defining calibration, we let $B_{L,\psi,\ell}(p) = \{R_L(u,p) ~:~ \psi(u) \not\in \gamma(p)\}$.
  Let $B_p = \inf B_{L,\psi,\ell}(p)$.
  By definition of calibration, we have $B_p > 0$.

@@

\paragraph{Part 2.}
Given Lemma \ref{lemma:fixed-p}, we might hope to obtain that, for any calibrated surrogate, there exists $\alpha = \sup_p \alpha_p$ such that $R_{\ell}(\psi(u),p') \leq \alpha R_L(u,p')$ for all $p'$.
However, we know this is false in general: not all surrogates yield linear surrogate regret transfer rates!
Indeed, for many surrogates, the supremum will be $+\infty$.

However, polyhedral surrogates have a special structure that allows us to show $\alpha$ is finite.
We first present several general lemmas, then the key implication from $L$ being polyhedral.
The general results are known \bo{cite}.
\begin{lemma} \label{lemma:refines}
  If $(L,\psi)$ indirectly elicits $\ell$, then $\Gamma = \prop{L}$ \emph{refines} $\gamma = \prop{\ell}$ in the sense that, for all $u \in \reals^d$, there exists $r \in \R$ such that $\Gamma_u \subseteq \Gamma_r$.
\end{lemma}
\begin{proof}
  For any $u$, let $r = \psi(u)$.
  By indirect elicitation, $u \in \Gamma(p) \implies r \in \gamma(p)$.
  Phrased differently, $\Gamma_u = \{p ~:~ u \in \Gamma(p)\} \subseteq \{p ~:~ r \in \gamma(p)\} = \gamma_r$.
\end{proof}
The next lemma states that surrogate regret is linear in $p$ on any fixed level set $\Gamma_{u^*}$ of $\Gamma = \prop{L}$.
By combining this fact with Lemma \ref{lemma:refines}, we obtain that target regret is linear on these level sets as well.
\begin{lemma} \label{lemma:linear-on-levelset}
  Suppose $(L,\psi)$ indirectly elicits $\ell$ and let $\Gamma = \prop{L}$.
  Then for any fixed $u,u^* \in \reals^d$ and $r \in \R$, the functions $R_L(u,\cdot)$ and $R_{\ell}(u,\cdot)$ are linear in their second arguments on $\Gamma_{u^*}$.
\end{lemma}
\begin{proof}
  Let, $p \in \Gamma_{u^*}$.
  By definition on $\Gamma_{u^*}$, $\risk{L}(p) = \inprod{p}{L(u^*)}$.
  So for fixed $u$,
    \[ R_L(u,p) = \inprod{p}{L(u)} - \inprod{p}{L(u^*)} = \inprod{p}{L(u) - L(u^*)} , \]
  a linear function of $p$.
  Next, by Lemma \ref{lemma:refines}, there exists $r^*$ such that $\Gamma_{u^*} \subseteq \gamma_{r^*}$.
  By the same argument, for fixed $r$, $R_{\ell}(r,p) = \inprod{p}{\ell(r) - \ell(r^*)}$, a linear function.
\end{proof}

We now arrive at the key property of polyhedral losses: \emph{although the prediction space is $\reals^d$, we can identify a finite set of predictions that characterize $L$.} \bo{wording?}
Here, a level set is \emph{full-dimensional} if it has affine dimension $|\Y|-1$, the same as the simplex $\simplex$.
\begin{lemma} \label{lemma:polyhedral-finite}
  If $L: \reals^d \to \reals_+^{\Y}$ is polyhedral, then $\Gamma = \prop{L}$ has a finite set of level sets that union to $\simplex$. Also, these level sets are polytopes.
\end{lemma}
\begin{proof}
  \bo{prove, cite, etc}
\end{proof}

We are now ready to restate and prove the main upper bound.
\begin{theorem}[Theorem \ref{thm:main-upper} restated]
  Let the surrogate loss $L: \reals^d \to \reals_+^{\Y}$ and link $\psi: \reals^d \to \R$ be consistent for the target loss $\ell: \R \to \reals_+^{\Y}$.
  If $L$ is polyhedral, then $(L,\psi)$ guarantee a linear regret transfer rate for $\ell$, i.e. there exists $\alpha > 0$ such that, for all $\D$ and all measurable $h: \X \to \reals^d$,
    \[ R_{\ell}(\psi \circ h ; \D) \leq \alpha R_L(h ; \D) . \]
\end{theorem}
\begin{proof}
  We first recall that by Fact \ref{fact:consistent-calibrated-elicits}, consistency implies that $(L,\psi)$ are calibrated for $\ell$ and that $(L,\psi)$ indirectly elicit $\ell$.
  Next, by Observation \ref{obs:transfer}, it suffices to show a linear \emph{conditional} regret transfer rate, i.e. for all $p \in \simplex$ and $u \in \reals^d$, we show $R_{\ell}(\psi(u),p) \leq \alpha R_L(u,p)$.

  By Lemma \ref{lemma:polyhedral-finite}, the polyhedral loss $L$ has a finite set $U \subset \reals^d$ of predictions such that (a) for each $u \in U$, the level set $\Gamma_u$ is a polytope, and (b) $\cup_{u \in U} \Gamma_u = \simplex$.
  Let $\mathcal{Q}_u \subseteq \simplex$ be the finite set of vertices of the polytope $\Gamma_u$, and let $\mathcal{Q} = \cup_{u \in U} \mathcal{Q}_u$.

  By Lemma \ref{lemma:fixed-p}, for each $p \in \mathcal{Q}$, there exists $\alpha_p \geq 0$ such that $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$ for all $u$.
  We choose
    \[ \alpha = \max_{p \in \mathcal{Q}} \alpha_p . \]
  To prove the conditional transfer rate, consider any $p \in \simplex$ and any $u \in \reals^d$.
  There exists $u \in U$ such that $p \in \Gamma_u$, so we can write $p$ as a convex combination of the vertices, i.e.
    \[ p = \sum_{p' \in \mathcal{Q}_u} \beta(p') p' \]
  for some probability distribution $\beta$.
  Because $\mathcal{Q}_u \subseteq \Gamma_u$, $R_L$ and $R_{\ell}$ are linear on it by Lemma \ref{lemma:linear-on-levelset}.
  So:
  \begin{align*}
    R_{\ell}(u,p)
    &=    R_{\ell}\left(\psi(u) ~,~ \sum_{p' \in \mathcal{Q}_u} \beta(p') p'\right)  \\
    &=    \sum_{p' \in \mathcal{Q}_u} \beta(p') R_{\ell}(\psi(u),p')  \\
    &\leq \sum_{p' \in \mathcal{Q}_u} \beta(p') \alpha_{p'} R_L(u,p')  \\
    &\leq \alpha \sum_{p' \in \mathcal{Q}_u} \beta(p') R_L(u,p')  \\
    &=    \alpha R_L(u, p) .
  \end{align*}
\end{proof}

\subsection{Part 1: fixed $p$}

We begin with the following key fact, a sort of ``pointwise'' conditional transfer bound.
\begin{proposition} \label{prop:fix-p}
  Suppose the surrogate and link $(L,\psi)$ are consistent for the target $\ell$.
  Then for any $p \in \simplex$, there exists $\alpha_p \geq 0$ such that, for all $u \in \reals^d$, $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$.
\end{proposition}

For intuition, observe that $R_{\ell}(\psi(u),p)$ takes on only finitely many values as we vary $p$, as the range $\R$ of $\psi$ is finite.
Therefore, it suffices to find a constant $C$ such that, if $R_{\ell}(\psi(u),p) > 0$, then $R_L(u,p) \geq C$.

\bo{include this proof sketch? Maybe not!}
\begin{enumerate}
  \item If $L$ is polyhedral, then the set of minimizers of $\inprod{p}{L(\cdot)}$ is a nice, convex, actually polyhedral subset of $\reals^d$.
  \item In addition, expected loss increases linearly with the distance of a prediction $u$ to this set.
  \item If $(L,\psi)$ is calibrated for $\ell$, then $\psi$ must be $\epsilon$-separated (Definition \bo{ref} below): all predictions with $\epsilon$ of the optimal set must link to an optimal target prediction in $\gamma(p)$.
  \item So conditional target regret is nonzero only for surrogate predictions that are $\epsilon$-far from the optimal set, which 
\end{enumerate}

>>>>

\section{Upper bound: polyhedral surrogates}
In this section, we develop the proof of Theorem \ref{thm:main-upper}.
At a high level, the proof has two parts:
\begin{enumerate}
  \item Fix $p \in \simplex$.
        Prove that $R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p)$ for some constant $\alpha_p$.
  \item Apply this result to a carefully chosen finite subset of $\simplex$. Argue that the maximum over the $\alpha_p$'s in this finite set gives the constant factor for the transfer rate.
\end{enumerate}
We view both parts as relatively surprising.
The first part will actually hold for any calibrated surrogate.
The second part relies on a crucial fact about polyhedral losses $L$: their properties $\Gamma$ cover the whole simplex with a finite number of polyhedral level sets.
In other words, although the prediction space $\reals^d$ is infinite, one can restrict to just a finite set of prediction points and always have a global minimizer of expected loss.
This is related to embeddings \bo{cite}.

\paragraph{Part 1.}
First, we establish a linear transfer rate for any fixed conditional distribution $p$.
This holds for any calibrated surrogate, not necessarily polyhedral.
\begin{lemma} \label{lemma:fixed-p}
  Let $\ell: \R \to \reals_+^{\Y}$ be a discrete target loss and suppose the surrogate $L: \reals^d \to \reals_+^{\Y}$ and link $\psi: \reals^d \to \R$ are calibrated for $\ell$.
  Then for any $p \in \simplex$, there exists $\alpha_p \geq 0$ such that, for all $u \in \reals^d$,
    \[ R_{\ell}(\psi(u),p) \leq \alpha_p R_L(u,p) . \]
\end{lemma}
\begin{proof}
  Fix $p \in \simplex$.
  Let $C_p = \max_{r \in \R} R_{\ell}(r,p)$.
  The maximum exists because $\ell$ is discrete, i.e. $\R$ is finite.
  Meanwhile, recall that, when defining calibration, we let $B_{L,\psi,\ell}(p) = \{R_L(u,p) ~:~ \psi(u) \not\in \gamma(p)\}$.
  Let $B_p = \inf B_{L,\psi,\ell}(p)$.
  By definition of calibration, we have $B_p > 0$.

@@

\paragraph{Part 2.}
Given Lemma \ref{lemma:fixed-p}, we might hope to obtain that, for any calibrated surrogate, there exists $\alpha = \sup_p \alpha_p$ such that $R_{\ell}(\psi(u),p') \leq \alpha R_L(u,p')$ for all $p'$.
However, we know this is false in general: not all surrogates yield linear transfer rates!
Indeed, for many surrogates, the supremum will be $+\infty$.

However, polyhedral surrogates have a special structure that allows us to show $\alpha$ is finite.
We first present several general lemmas, then the key implication from $L$ being polyhedral.
The general results are known \bo{cite}.

\bo{For these lemmas, I tried to state no more than we need.}
\begin{lemma} \label{lemma:refines}
  If $(L,\psi)$ indirectly elicits $\ell$, then $\Gamma = \prop{L}$ \emph{refines} $\gamma = \prop{\ell}$ in the sense that, for all $u \in \reals^d$, there exists $r \in \R$ such that $\Gamma_u \subseteq \gamma_r$.
\end{lemma}
\begin{proof}
  For any $u$, let $r = \psi(u)$.
  By indirect elicitation, $u \in \Gamma(p) \implies r \in \gamma(p)$.
  Phrased differently, $\Gamma_u = \{p ~:~ u \in \Gamma(p)\} \subseteq \{p ~:~ r \in \gamma(p)\} = \gamma_r$.
\end{proof}
The next lemma states that surrogate regret is linear in $p$ on any fixed level set $\Gamma_{u^*}$ of $\Gamma = \prop{L}$.
By combining this fact with Lemma \ref{lemma:refines}, we obtain that target regret is linear on these level sets as well.
\begin{lemma} \label{lemma:linear-on-levelset}
  Suppose $(L,\psi)$ indirectly elicits $\ell$ and let $\Gamma = \prop{L}$.
  Then for any fixed $u,u^* \in \reals^d$ and $r \in \R$, the functions $R_L(u,\cdot)$ and $R_{\ell}(u,\cdot)$ are linear in their second arguments on $\Gamma_{u^*}$.
\end{lemma}
\begin{proof}
  Let $u^* \in \reals^d$ and $p \in \Gamma_{u^*}$.
  By definition, for all $p \in \Gamma_{u^*}$, $\risk{L}(p) = \inprod{p}{L(u^*)}$.
  So for fixed $u$,
    \[ R_L(u,p) = \inprod{p}{L(u)} - \inprod{p}{L(u^*)} = \inprod{p}{L(u) - L(u^*)} , \]
  a linear function of $p$.
  Next, by Lemma \ref{lemma:refines}, there exists $r^*$ such that $\Gamma_{u^*} \subseteq \gamma_{r^*}$.
  By the same argument, for fixed $r$, $R_{\ell}(r,p) = \inprod{p}{\ell(r) - \ell(r^*)}$, a linear function.
\end{proof}

We now arrive at the key property of polyhedral losses.
\begin{lemma} \label{lemma:polyhedral-finite}
  If $L: \reals^d \to \reals_+^{\Y}$ is polyhedral, then $\Gamma = \prop{L}$ has a finite set of level sets that union to $\simplex$. Also, these level sets are polytopes.
\end{lemma}
\begin{proof}
  \bo{prove, cite, etc}
\end{proof}

We are now ready to restate and prove the main upper bound.
\begin{theorem}[Theorem \ref{thm:main-upper} restated]
  Suppose the surrogate loss $L: \reals^d \to \reals_+^{\Y}$ and link $\psi: \reals^d \to \R$ are consistent for the target loss $\ell: \R \to \reals_+^{\Y}$.
  If $L$ is polyhedral, then $(L,\psi)$ guarantee a linear regret transfer rate for $\ell$, i.e. there exists $\alpha > 0$ such that, for all $\D$ and all measurable $h: \X \to \reals^d$,
    \[ R_{\ell}(\psi \circ h ; \D) \leq \alpha R_L(h ; \D) . \]
\end{theorem}
\begin{proof}
  We first recall that by Fact \ref{fact:consistent-calibrated-elicits}, consistency implies that $(L,\psi)$ are calibrated for $\ell$ and that $(L,\psi)$ indirectly elicit $\ell$.
  Next, by Observation \ref{obs:transfer}, it suffices to show a linear \emph{conditional} regret transfer rate, i.e. for all $p \in \simplex$ and $u \in \reals^d$, we show $R_{\ell}(\psi(u),p) \leq \alpha R_L(u,p)$.

  By Lemma \ref{lemma:polyhedral-finite}, the polyhedral loss $L$ has a finite set $U \subset \reals^d$ of predictions such that (a) for each $u \in U$, the level set $\Gamma_u$ is a polytope, and (b) $\cup_{u \in U} \Gamma_u = \simplex$.
  Let $\mathcal{Q}_u \subset \simplex$ be the finite set of vertices of the polytope $\Gamma_u$, and define the finite set $\mathcal{Q} = \cup_{u \in U} \mathcal{Q}_u$.

  By Lemma \ref{lemma:fixed-p}, for each $q \in \mathcal{Q}$, there exists $\alpha_q \geq 0$ such that $R_{\ell}(\psi(u),q) \leq \alpha_p R_L(u,q)$ for all $u$.
  We choose
    \[ \alpha = \max_{q \in \mathcal{Q}} \alpha_q . \]
  To prove the conditional transfer rate, consider any $p \in \simplex$ and any $u \in \reals^d$.
  There exists $u \in U$ such that $p \in \Gamma_u$, a polytope.
  So we can write $p$ as a convex combination of its vertices, i.e.
    \[ p = \sum_{q \in \mathcal{Q}_u} \beta(q) q \]
  for some probability distribution $\beta$.
  Because $\mathcal{Q}_u \subseteq \Gamma_u$, $R_L$ and $R_{\ell}$ are linear on it by Lemma \ref{lemma:linear-on-levelset}.
  So:
  \begin{align*}
    R_{\ell}(u,p)
    &=    R_{\ell}\left(\psi(u) ~,~ \sum_{q \in \mathcal{Q}_u} \beta(q) q\right)  \\
    &=    \sum_{q \in \mathcal{Q}_u} \beta(q) R_{\ell}(\psi(u),q)  \\
    &\leq \sum_{q \in \mathcal{Q}_u} \beta(q) \alpha_{q} R_L(u,q)  \\
    &\leq \alpha \sum_{q \in \mathcal{Q}_u} \beta(q) R_L(u,q)  \\
    &=    \alpha R_L(u, p) .
  \end{align*}
\end{proof}

\section{Lower bound}

\section{Determining the constant for polyhedral losses}

\bo{Stuff about Hoffman constants, etc}
